Hydra Battles and Cie

### 2021-01-20 doc/chapter-primrec.tex
corrections to the Ackermann doc

\index{Projects}
\begin{project}
Prove formally that the Ackermann function is \emph{not} primitive recursive.
You may find a definition of the Ackermann function and a few  lemmas in
 \href{../theories/ordinals/Prelude/Iterates.html}%
{ordinals.Prelude.Iterates}.

\end{project}

>>>>

\index{Exercises}
\begin{exercise}
Show that the function \texttt{min: naryFunc\,2} is primitive recursive.
\end{exercise}

\index{Projects}
\begin{project}
Prove formally that the Ackermann function is \emph{not} primitive recursive.

\begin{align}
A\,0\,m&=m+1\\
A\,(n+1)\,0 &= A\,n\,1\\
A\,(n+1)\,(m+1)&= A\,n\,(A\,(n+1)\,m)
\end{align}

You may find also  definition of the Ackermann function and a few  lemmas in
 \href{../theories/ordinals/Prelude/Iterates.html}%
{ordinals.Prelude.Iterates}.

\textbf{Hint:} the reader may follow the plan suggested by Bruno Salvy (examen given at the French \'Ecole Polytechnique) (available at
\href{http://www.enseignement.polytechnique.fr/informatique/INF412/uploads/Main/pc-primrec-sujet2014.pdf}{this site}).

\begin{enumerate}
\item Show that, for any $n$, the function $A\,n$ is primitive recursive.
\item Show that for all $n$ and $m$, $A\,n\,m > m$ and  that the functions $A\,n$ are strictly monotonous.
Show also that, for any $n$ and $m$, $A\,n\,m \leq A\,(n+1)\,m$.
         Finally, show that for any $n$, $m$ and $k$, 
$A\,k\,(A\,m\,n)\leq A\,(2+\textrm{max}(k,m))\,n$.
\item Show that for any primitive recursive function $f$ of $k$ arguments,
for any $x_1,\dots\,x_k$, $f(x_1,\dots,x_k)\leq A\,n\,(x_1+\dots+x_k)$.
\item Show that the Ackermann function is not primitive recursive.

\end{enumerate}

\begin{remark}
Note that the Ackermann function is a counter-example to the (false) statement:
\begin{quote}
  ``Let $f$ be a function of type \texttt{naryFunc\,2}. If, for any $n$, the fonction $f(n)$ is primitive recursive , then f is primitive recursive.''
\end{quote}
\end{remark}
\end{project}
